### 2021-04-14 submission/manuscript.Rmd
Add M&I PDA writing group to acknowledgements
Add results text

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared PEG-treated mice to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^3^ *C. difficile* 630 spores. All treatments rendered mice susceptible to *C. difficile* colonization (Fig. 1C), however PEG-treated mice remained colonized at a high level through 30 days post-infection. In contrast, the clindamycin-treated mice that cleared *C. difficile* within 10 days post-infection. Surprisingly, mice were still susceptible to *C. difficile* infection after 10-days of recovery from PEG treatment although *C. difficile* was not detectable in most of the group in the initial 5 days post-infection (Fig. 1C). From 9 days post-infection onward, the median *C. difficile* stabilized for the 5-day PEG plus 10-day recovery group of mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible for up to 10 days post PEG treatment.
**5-day laxative treatment differentially disrupts the fecal microbiota compared to clindamycin treatment.** Given 
**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given

* Figure 1. 5-day PEG treatment prolongs susceptibility and mice become persistently colonized with *C. difficile*.

@@

## Acknowledgements
We thank members of the Schloss lab for feedback on planning the experiments and data presentation. We also thank Andrew Henry for help with media preparation and bacterial culture. We also thank the Unit for Laboratory Animal Medicine at the University of Michigan for maintaining our mouse colony and providing the institutional support for our mouse experiments. Finally, we thank Kwi Kim, Austin Campbell, and Kimberly Vendrov for their help in maintaining the Schloss lab's anaerobic chamber. This work was supported by the National Institutes of Health (U01AI124255). ST was supported by the Michigan Institute for Clincial and Health Research Postdoctoral Translation Scholars Program (UL1TR002240 from the National Center for Advancing Translational Sciences).

>>>>

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared PEG-treated mice to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^3^ *C. difficile* 630 spores. All treatments rendered mice susceptible to *C. difficile* colonization (Fig. 1C), however PEG-treated mice remained colonized at a high level through 30 days post-infection. In contrast, the clindamycin-treated mice that cleared *C. difficile* within 10 days post-infection. Surprisingly, mice were still susceptible to *C. difficile* infection after 10-days of recovery from PEG treatment although *C. difficile* was not detectable in most of the group in the initial 5 days post-infection (Fig. 1C). From 9 days post-infection onward, the median *C. difficile* stabilized for the 5-day PEG plus 10-day recovery group of mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible for up to 10 days post PEG treatment.

``` {r 5-day PEG microbiota}
f_d_permanova <- read_tsv(here("data/process/5_days_PEG_permanova_stools.tsv"))
f_d_top_2_contrib <- f_d_permanova %>% arrange(desc(R2)) %>%
  filter(!effects == "Total") %>% #Remove total since it is not one of the variables
  top_n(2, R2) %>% #Select top 2 variables based on R2
  select(effects, R2, `Pr(>F)`) #Select relevant column names

f_d_top_contrib= (f_d_top_2_contrib[1,1])
f_d_sec_contrib = (f_d_top_2_contrib[2,1])
f_d_top_R2_sum = f_d_top_2_contrib %>% 
  select(R2) %>% sum
f_d_top_p <- f_d_top_2_contrib %>% 
  select(`Pr(>F)`) %>% max # Both p values are the same, will be rounded in text
f_d_top_R2 = f_d_top_2_contrib[1,2]
f_d_sec_R2 = f_d_top_2_contrib[2,2]
```
**5-day laxative treatment differentially disrupts the fecal microbiota compared to clindamycin treatment.** Since laxatives and clindamycin have previously been shown to disrupt the murine microbiota (ref), we hypothesized the different *C. difficile* colonization dynamics between mice treated with laxatives versus clindamycin were due to their differential effects on the microbiota. We profiled the stool microbiota over time by sequencing the V4 region of the 16S rRNA gene to compare changes across treatment groups. We found time and treatment group explained half of the observed variation between fecal communities (PERMANOVA combined R^2^ = `r f_d_top_R2_sum`, *P* `r format.pval(f_d_top_p, eps = 0.001)`) with the remaining variation explained by interactions between treatment group and other experimental variables including time, cage effects, and sequencing preparation (Data Set S1, Sheet X). The impact of cage effects has been well documented due to copraphagy and is expected.

**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given previous work has shown that PEG treatment disrupts the mucus layer, we decided to examine the impact of the laxative on the mucosal microbiota and infection severity by evaluating histopathology.

@@

## Acknowledgements
We thank members of the Schloss lab for feedback on planning the experiments and data presentation. We thank Andrew Henry for help with media preparation and bacterial culture and the Microbiology and Immunology department's postdoc association writing group members for their feedback on manuscript drafts. We also thank the Unit for Laboratory Animal Medicine at the University of Michigan for maintaining our mouse colony and providing the institutional support for our mouse experiments. Finally, we thank Kwi Kim, Austin Campbell, and Kimberly Vendrov for their help in maintaining the Schloss lab's anaerobic chamber. This work was supported by the National Institutes of Health (U01AI124255). ST was supported by the Michigan Institute for Clincial and Health Research Postdoctoral Translation Scholars Program (UL1TR002240 from the National Center for Advancing Translational Sciences).

### 2021-04-14 submission/manuscript.tex
Add results text

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing
\emph{C. difficile} within 10 days post-infection (9, 19). All
PEG-treated mice were administered a 15\% PEG solution in the drinking
water for 5-days, one group was also treated with clindamycin, and one
group was allowed to recover for 10 days prior to challenge (Fig. 1A).
PEG treatment resulted in weight loss in all 3 groups of PEG-treated
mice, with the greatest change in weight observed on the fifth day of
PEG treatment (Fig. 1B). After either PEG, clindamycin, or PEG and
clindamycin treatment all mice were challenged with
10\textsuperscript{3} \emph{C. difficile} 630 spores. All treatments
rendered mice susceptible to \emph{C. difficile} colonization (Fig. 1C),
however PEG-treated mice remained colonized at a high level through 30
days post-infection. In contrast, the clindamycin-treated mice that
cleared \emph{C. difficile} within 10 days post-infection. Surprisingly,
mice were still susceptible to \emph{C. difficile} infection after
10-days of recovery from PEG treatment although \emph{C. difficile} was
not detectable in most of the group in the initial 5 days post-infection
(Fig. 1C). From 9 days post-infection onward, the median \emph{C.
difficile} stabilized for the 5-day PEG plus 10-day recovery group of
mice and remained high through 30 days post-infection (Fig. 1C). Thus,
osmotic laxative treatment alone was sufficient to render mice
susceptible to prolonged \emph{C. difficile} colonization and
PEG-treated mice remained susceptible for up to 10 days post PEG
treatment. \textbf{5-day laxative treatment differentially disrupts the
fecal microbiota compared to clindamycin treatment.} Given \textbf{5-day
laxative treatment does not promote more severe CDIs despite altering
the mucosal microbiota.} Given

\begin{itemize}
\item
  Figure 1. 5-day PEG treatment prolongs susceptibility and mice become
  persistently colonized with \emph{C. difficile}.
\item
  Figure 2. 5-day PEG treatment disrupts the stool microbiota for a
  longer amount of time compared to clindamycin-treated mice.
\item
  Figure S1. 5-day PEG treatment plus 10-day recovery mice microbiota
  dynamics post-infection.
\item
  Figure 3. 5-day PEG treatment does not result in more severe CDIs,
  although mucosal microbiota is altered.
\item
  Figure 4. 1-day PEG treatment renders mice susceptible to transient
  \emph{C. difficile} colonization.
\item
  Figure 5. 1-day PEG treatment post \emph{C. difficile} challenge
  prolongs colonization regardless of whether an FMT is also
  administered.
\item
  Figure 6. Specific microbiota features associated with prolonged
  \emph{C. difficile} colonization in PEG treated mice.
\item
  Figure S2. Specific OTUs associated with clearance that are mostly
  absent in mice with prolonged \emph{C. difficile} colonization.

@@

We thank members of the Schloss lab for feedback on planning the
experiments and data presentation. We also thank Andrew Henry for help
with media preparation and bacterial culture. We also thank the Unit for
Laboratory Animal Medicine at the University of Michigan for maintaining
our mouse colony and providing the institutional support for our mouse
experiments. Finally, we thank Kwi Kim, Austin Campbell, and Kimberly
Vendrov for their help in maintaining the Schloss lab's anaerobic
chamber. This work was supported by the National Institutes of Health
(U01AI124255). ST was supported by the Michigan Institute for Clincial
and Health Research Postdoctoral Translation Scholars Program
(UL1TR002240 from the National Center for Advancing Translational
Sciences).

>>>>

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared PEG-treated mice to our standard 10
mg/kg clindamycin treatment, which temporarily renders the mice
susceptible to \emph{C. difficile}, with mice typically clearing
\emph{C. difficile} within 10 days post-infection (9, 19). All
PEG-treated mice were administered a 15\% PEG solution in the drinking
water for 5-days, one group was also treated with clindamycin, and one
group was allowed to recover for 10 days prior to challenge (Fig. 1A).
PEG treatment resulted in weight loss in all 3 groups of PEG-treated
mice, with the greatest change in weight observed on the fifth day of
PEG treatment (Fig. 1B). After either PEG, clindamycin, or PEG and
clindamycin treatment all mice were challenged with
10\textsuperscript{3} \emph{C. difficile} 630 spores. All treatments
rendered mice susceptible to \emph{C. difficile} colonization (Fig. 1C),
however PEG-treated mice remained colonized at a high level through 30
days post-infection. In contrast, the clindamycin-treated mice that
cleared \emph{C. difficile} within 10 days post-infection. Surprisingly,
mice were still susceptible to \emph{C. difficile} infection after
10-days of recovery from PEG treatment although \emph{C. difficile} was
not detectable in most of the group in the initial 5 days post-infection
(Fig. 1C). From 9 days post-infection onward, the median \emph{C.
difficile} stabilized for the 5-day PEG plus 10-day recovery group of
mice and remained high through 30 days post-infection (Fig. 1C). Thus,
osmotic laxative treatment alone was sufficient to render mice
susceptible to prolonged \emph{C. difficile} colonization and
PEG-treated mice remained susceptible for up to 10 days post PEG
treatment.

\textbf{5-day laxative treatment differentially disrupts the fecal
microbiota compared to clindamycin treatment.} Since laxatives and
clindamycin have previously been shown to disrupt the murine microbiota
(ref), we hypothesized the different \emph{C. difficile} colonization
dynamics between mice treated with laxatives versus clindamycin were due
to their differential effects on the microbiota. We profiled the stool
microbiota over time by sequencing the V4 region of the 16S rRNA gene to
compare changes across treatment groups. We found time and treatment
group explained half of the observed variation between fecal communities
(PERMANOVA combined R\textsuperscript{2} = 0.50, \emph{P} \textless{}
0.001).

\textbf{5-day laxative treatment does not promote more severe CDIs
despite altering the mucosal microbiota.} Given previous work has shown
that PEG treatment disrupts the mucus layer, we decided to examine the
impact of the laxative on the mucosal microbiota and infection severity
by evaluating histopathology.

\begin{itemize}
\item
  Figure 2. 5-day PEG treatment disrupts the stool microbiota for a
  longer amount of time compared to clindamycin-treated mice.
\item
  Figure S1. 5-day PEG treatment plus 10-day recovery mice microbiota
  dynamics post-infection.
\item
  Figure 3. 5-day PEG treatment does not result in more severe CDIs,
  although mucosal microbiota is altered.
\item
  Figure 4. 1-day PEG treatment renders mice susceptible to transient
  \emph{C. difficile} colonization.
\item
  Figure 5. 1-day PEG treatment post \emph{C. difficile} challenge
  prolongs colonization regardless of whether an FMT is also
  administered.
\item
  Figure 6. Specific microbiota features associated with prolonged
  \emph{C. difficile} colonization in PEG treated mice.
\item
  Figure S2. Specific OTUs associated with clearance that are mostly
  absent in mice with prolonged \emph{C. difficile} colonization.

@@

We thank members of the Schloss lab for feedback on planning the
experiments and data presentation. We thank Andrew Henry for help with
media preparation and bacterial culture and the Microbiology and
Immunology department's postdoc association writing group members for
their feedback on manuscript drafts. We also thank the Unit for
Laboratory Animal Medicine at the University of Michigan for maintaining
our mouse colony and providing the institutional support for our mouse
experiments. Finally, we thank Kwi Kim, Austin Campbell, and Kimberly
Vendrov for their help in maintaining the Schloss lab's anaerobic
chamber. This work was supported by the National Institutes of Health
(U01AI124255). ST was supported by the Michigan Institute for Clincial
and Health Research Postdoctoral Translation Scholars Program
(UL1TR002240 from the National Center for Advancing Translational
Sciences).
